\subsection{Interconnect-Manager Hardware Custom Instruction Library}

\begin{itemize}
\item \textbf{Broadcast Intracellular Message}
Send a message to all SignalGP instances within the cell, excluding self.
\item \textbf{Put Internal Membrane Bringer}
Place a tag in the internal membrane that, with default arguments, admits incoming internal messages it matches with.
\item \textbf{Put Internal Membrane Blocker}
Place a tag in the internal membrane that, with default arguments, blocks incoming internal messages it matches with.
\item \textbf{Put External Membrane Bringer}
Place a tag in the external membrane that, with default arguments, admits incoming external messages it matches with.
\item \textbf{Put External Membrane Blocker}
Place a tag in the external membrane that, with default arguments, blocks incoming external messages it matches with.
\item \textbf{Broadcast Intercellular Message}
Send a message to all cellular neighbors.
\item \textbf{Send Interconnect Message}
Send a message to the interconnect-manager hardware component of all cells the executing cell has established an interconnect to.
\item \textbf{Broadcast Interconnect Message}
Send a message to all the neighbor-manager and interconnect-manager hardware components of all cells the executing cell has established an interconnect to.
All interconnect broadcast messages incoming to a cell during a delivery interval are de-duplicated.
(If several identical messages are broadcast to a cell via interconnect during a delivery interval, only one message is delivered to each hardware component of the cell.)
\item \textbf{Send Reverse Interconnect Message}
Send a message to the interconnect-manager hardware component of all cells that have established an interconnect to the executing cell.
\item \textbf{Broadcast Reverse Interconnect Message}
Send a message to all the neighbor-manager and interconnect-manager hardware components of all cells that have established interconnects to the executing cell.
All interconnect broadcast messages incoming to a cell during a delivery interval are de-duplicated.
(If several identical messages are broadcast to a cell via interconnect during a delivery interval, only one message is delivered to each hardware component of the cell.)
\item \textbf{Share Resource}
Send a proportion of the cell's stockpiled resource to all cells the executing cell has established an interconnect to.
The proportion utgoing resource is split among cells the executing cell has established an interconnect to.
If no outgoing interconnect has been established, no operation takes place.
One instruction defaults to sending a large proportion of available resource (50\%) to the neighboring cell.
A second instruction defaults to sending a small proportion of available resource (5\%) to the neighboring cell.
The proportion of available resource can be adjusted by a register-based argument.
\item \textbf{Set Stockpile Sharing Reserve}
Designate a quantity of stockpiled resource as ineligible for sharing.
The amount may be modified by a register-based argument.
\item \textbf{Clear Stockpile Sharing Reserve}
Designate all stockpiled resource as eligible for sharing.
\item \textbf{Restrict Outgoing Shared Resource}
Reduce outgoing sharing efficacy.
Unsent resource is retained by the sending cell (with no resource lost).
The fraction reduced is determined by a register-based argument.
\item \textbf{Restrict Incoming Shared Resource}
Reduce incoming sharing efficacy.
Declined resource is retained by the sending cell (with no resource lost).
The fraction reduced is determined by a register-based argument.
\item \textbf{Reproduce}
Attempt to spawn a child cell in a particular direction, paid for out of the parent cell's resource stockpile.
If sufficient resource is not available in the cell's stockpile, no resource is action is taken.
Variants of this instruction are defined for each channel ID inheritance level: from endowing the daughter cell with the parental channel IDs across all levels, to endowing the daughter cell with a new level-one channel ID but the parent's level-two channel ID, to endowing the daughter cell with all-new channel IDs.
If a channel generation counter limit has been reached, reproduction is simply attempted at the next highest level; even with channel generation counters maxed out, cells may generate offspring with all-new channel IDs.
\item \textbf{Mark Daughter Imprint}
Depending on argument values, register or clear a tag imprint to apply to daughter cells.
\item \textbf{Pause Reproduction}
Pause cellular reproduction in a single direction for the remainder of the current update and for the entire next update.
Variants of this instruction pause reproduction at a certain wave/channel-signaling level or across all channel ID inheritance levels.
\item \textbf{Set Stockpile Reproduction Reserve}
Designate a quantity of stockpiled resource as ineligible for use to reproduce.
The amount may be modified by a register-based argument.
\item \textbf{Clear Stockpile Sharing Reserve}
Designate all stockpiled resource as eligible for use to reproduce.
\item \textbf{Apoptosis}
The cell is killed at the end of the current update.
\item \textbf{Designate Heir} A dying cell's own stockpile is split evenly among neighboring cells that are designated at the time of death.
On apoptosis, 50\% of the reproduction cost to establish a cell is also split between designated neighboring cells.
These instructions mark or un-mark a neighbor as a heir.
\item \textbf{Increase Channel Generation Counter}
Increases the cell's channel generation counter.
The amount the cell's generation counter is increased by can be adjusted by register-based argument.
\item \textbf{Query Own Stockpile}
Sets a designated register to the amount of resource present in the cell's stockpile.
\item \textbf{Query Own Channel Generation Counter}
This instruction sets a designated register to the value of the cell's channel generation counter.
A variant of this instruction is provided for each wave/channel-signaling level.
\item \textbf{Query Incoming Connections}
This instruction writes the current number of incoming connections and the current mean length of incoming connections into two separate genetically-determined registers.
\item \textbf{Query Outgoing Connections}
This instruction writes the current number of outgoing connections and the current mean length of outgoing connections into two separate genetically-determined registers.
\item \textbf{Add Fledgling Connection}
If the number of outgoing connections is less than both the system-defined connection count cap and an argument-defined connection count cap, start a new fledgling interconnect.
\item \textbf{Add Fledgling Search Query}
Add a tag query that, depending on argument value, either promotes or represses the cell's fledgling interconnects to mature into established interconnects with other cells containing tag signals it matches with.
\item \textbf{Add Fledgling Search Signal}
Add a tag signal that, depending on argument value, either promotes or represses other cell's fledgling interconnects with tag queries it matches with to mature into an established interconnect.
\item \textbf{Prune Outgoing Connection}
Remove an outgoing established interconnect.
\item \textbf{Prune Outgoing Connection}
Remove an incoming established interconnect.
\item \textbf{Set Fledgling Connection Aging Parameter}
Set a parameter that controls the baseline rate at which fledgling interconnects mature towards becoming an established interconnect.
\item \textbf{Set Fledgling Connection Search Exploitation Parameter}
Set a parameter that controls the baseline rate at which less-successful fledgling interconnect probes reset to more-successful fledgling interconnect  probes.
\item \textbf{Set Fledgling Connection Threshold Parameter}
Set a parameter that controls the threshold at which fledgling interconnects mature into established interconnects.
\item \textbf{Set Fledgling Connection Query Sensitivity Parameter}
Set a parameter that controls the fledgling interconnect's sensitivity to favorable and unfavorable tagged query/signal matches.
\end{itemize}


