\section{Interconnect-Manager Hardware Environmental Cue Library}

Environmental cues were triggered every eight updates.

\begin{itemize}
\item \textbf{On Update}
This event is triggered every eight updates.
\item \textbf{Imprinted Tag}
If an imprinted tag was set by the cell's parent, this event is triggered every eight updates.
\item \textbf{Just Born}
This event is triggered once after a cell is born.
\item \textbf{Connection Length}
Separate events are triggered if a cell's mean outgoing interconnect length falls between 0 and 4 cell widths, 2 and 6 cell widths, 4 and 8 cell widths, 6 and 10 cell widths, or above 8 cell widths.
\item \textbf{Channel Generation Counter Is Unexpired}
This event is triggered if a SignalGP instance's cell's channel generation counter has not yet reached the expiration threshold.
A different event is provided for each resource wave/channel-signaling level.
\item \textbf{Channel Generation Counter Is Expiring}
This event is triggered if a SignalGP instance's cell's channel generation counter has not yet reached the threshold where somatic propagation capacity, but not resource accumulation capacity, is lost.
A different event is provided for each resource wave/channel-signaling level.
\item \textbf{Channel Generation Counter Is Expired}
This event is triggered if a SignalGP instance's cell's channel generation counter has not yet reached the threshold where both somatic propagation capacity and resource accumulation capacity are lost.
A different event is provided for each resource wave/channel-signaling level.
\item \textbf{No-reward Resource Activation}
This event is triggered if a SignalGP instance's cell experiences a resource collection activation where no resource reward is achieved (e.g., the cell lies extent of the resource wave).
A different event is provided for each resource wave/channel-signaling level.
\end{itemize}

