\section{Results and Discussion} \label{sec:results}

\input{tex/proof1.tex}

\input{tex/proof2.tex}

\input{tex/proof3.tex}

\input{tex/proof4.tex}

% \input{fig/grid_progression.tex}

\subsection{Case Study: Interconnect Resource Sharing}

\input{fig/case_study_1042.tex}

This case study was drawn from epoch 24 of batch 42 of the initial set of evolutionary runs.
We initially considered it for further study due to the presence of widespread over-interconnect resource sharing.
After preliminary knockout experiments confirmed the adaptive significance of both over-interconnect resource-sharing and over-interconnect messaging, we set aside the strain for a case study.
The evolutionary history preceding this case study consumed approximately 96 hours of wall-clock time and 736 compute-core hours.
Approximately 29,873,088 simulation updates and 41,439 cellular generations elapsed.
You can view the strain this case study characterizes in a live in-browser simulation at \url{https://mmore500.com/hopto/8}

% 100%|████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████| 25/25 [00:00<00:00, 292.67it/s]
% Output saved to title=mastergenerations+_data_hathash_hash=dc05faa593102bf1+_script_fullcat_hash=5e6b7a64514cbb09+_source_hash=aa98eba-clean+ext=.csv
% level 0
% Generations Elapsed mean 41438.728819430224
% Generations Elapsed std nan
% Generations Elapsed min 41438.728819430224
% Generations Elapsed max 41438.728819430224
% level 1
% Generations Elapsed mean 24466.75578134247
% Generations Elapsed std nan
% Generations Elapsed min 24466.75578134247
% Generations Elapsed max 24466.75578134247
% level 2
% Generations Elapsed mean 5130.206547616208
% Generations Elapsed std nan
% Generations Elapsed min 5130.206547616208
% Generations Elapsed max 5130.206547616208

% EXPIRATION_UPDATE 361120
% EXPIRATION_UPDATE 336640
% EXPIRATION_UPDATE 439872
% EXPIRATION_UPDATE 247360
% EXPIRATION_UPDATE 255456
% EXPIRATION_UPDATE 239904
% EXPIRATION_UPDATE 230144
% EXPIRATION_UPDATE 119168
% EXPIRATION_UPDATE 232736
% EXPIRATION_UPDATE 238400
% EXPIRATION_UPDATE 270912
% EXPIRATION_UPDATE 125856
% EXPIRATION_UPDATE 204256
% EXPIRATION_UPDATE 242368
% EXPIRATION_UPDATE 190432
% EXPIRATION_UPDATE 270880
% EXPIRATION_UPDATE 244544
% EXPIRATION_UPDATE 216064
% EXPIRATION_UPDATE 282688
% EXPIRATION_UPDATE 254048
% EXPIRATION_UPDATE 309120
% EXPIRATION_UPDATE 279904
% EXPIRATION_UPDATE 342848
% EXPIRATION_UPDATE 289248
% EXPIRATION_UPDATE 237664
% EXPIRATION_UPDATE 166016
% EXPIRATION_UPDATE 251520
% EXPIRATION_UPDATE 207424
% EXPIRATION_UPDATE 260064
% EXPIRATION_UPDATE 248480
% EXPIRATION_UPDATE 301312
% EXPIRATION_UPDATE 225088
% EXPIRATION_UPDATE 347072
% EXPIRATION_UPDATE 321120
% EXPIRATION_UPDATE 344448
% EXPIRATION_UPDATE 388256
% EXPIRATION_UPDATE 382336
% EXPIRATION_UPDATE 454560
% EXPIRATION_UPDATE 413312
% EXPIRATION_UPDATE 404512
% EXPIRATION_UPDATE 437920
% EXPIRATION_UPDATE 402560
% EXPIRATION_UPDATE 435552
% EXPIRATION_UPDATE 420384
% EXPIRATION_UPDATE 366752
% EXPIRATION_UPDATE 401440
% EXPIRATION_UPDATE 318752
% EXPIRATION_UPDATE 395936
% EXPIRATION_UPDATE 176608
% EXPIRATION_UPDATE 194656
% EXPIRATION_UPDATE 228928
% EXPIRATION_UPDATE 115776
% EXPIRATION_UPDATE 272000
% EXPIRATION_UPDATE 282176
% EXPIRATION_UPDATE 223232
% EXPIRATION_UPDATE 176320
% EXPIRATION_UPDATE 334080
% EXPIRATION_UPDATE 359328
% EXPIRATION_UPDATE 404000
% EXPIRATION_UPDATE 277824
% EXPIRATION_UPDATE 206592
% EXPIRATION_UPDATE 189024
% EXPIRATION_UPDATE 216736
% EXPIRATION_UPDATE 166624
% EXPIRATION_UPDATE 194560
% EXPIRATION_UPDATE 198336
% EXPIRATION_UPDATE 182080
% EXPIRATION_UPDATE 173824
% EXPIRATION_UPDATE 309440
% EXPIRATION_UPDATE 281888
% EXPIRATION_UPDATE 345280
% EXPIRATION_UPDATE 326144
% EXPIRATION_UPDATE 259680
% EXPIRATION_UPDATE 292096
% EXPIRATION_UPDATE 205056
% EXPIRATION_UPDATE 216672
% EXPIRATION_UPDATE 355360
% EXPIRATION_UPDATE 401824
% EXPIRATION_UPDATE 388768
% EXPIRATION_UPDATE 434080
% EXPIRATION_UPDATE 280480
% EXPIRATION_UPDATE 307104
% EXPIRATION_UPDATE 350592
% EXPIRATION_UPDATE 279424
% EXPIRATION_UPDATE 424000
% EXPIRATION_UPDATE 404160
% EXPIRATION_UPDATE 412640
% EXPIRATION_UPDATE 376768
% EXPIRATION_UPDATE 325216
% EXPIRATION_UPDATE 366816
% EXPIRATION_UPDATE 380384
% EXPIRATION_UPDATE 361664
% EXPIRATION_UPDATE 371488
% EXPIRATION_UPDATE 398400
% EXPIRATION_UPDATE 374816
% EXPIRATION_UPDATE 368416
% EXPIRATION_UPDATE 329600
% EXPIRATION_UPDATE 302016
% EXPIRATION_UPDATE 328832
% EXPIRATION_UPDATE 384832


Our first step was to evaluate whether intercellular nature of over-interconnect messaging and resource sharing contributed to the strain's fitness.
(It is possible that messaging and/or resource sharing behaviors might generate stimuli on the recipient or side-effects on the sender that have adaptive consequences whether or not the sender and recipient are distinct cells;
in such a scenario, cells would be just as well off sending messages and/or resource to themselves.)
We performed several competition experiments between between the wild-type strain and variants where interconnect messaging and resource sharing was altered to be intracellular instead of intercellular.
At the end of competition experiments, we evaluated the relative abundances of wild-type and variant strains.
In the first variant strain we tested, all outgoing over-interconnect messages were delivered to the sending cell.
In 16 out of 16 one-hour competition runs that were seeded half-and-half with the wild-type and variant strains, the wild-type strain drove the variant strain to extinction (one-tailed binomial test; $p < 0.0001$; 290 S.D 17 cell gens elapsed).
We observed a similar outcome with a second variant strain where all outgoing over-interconnect resource sharing was rerouted back to the sending cell (14/16 variant strain extinctions; 16/16 wild-type prevalence; 289 S.D. 25 cell gens elapsed).
Finally, a third variant strain where both over-interconnect messaging and over-interconnect resource sharing were returned to the sending cell exhibited the same outcome (16/16 variant strain extinctions; 300 S.D. 23 cell gens elapsed).
The intercellular natures of both over-interconnect messaging and resource sharing appear essential to fitness.

Next, we took a closer look at the evolved cellular mechanisms controlling over-interconnect messaging and resource sharing.
We monitored hardware execution of the wild-type strain in a monoculture population to detect which signals, messages, and fork/call instructions activated each SignalGP module.
We manually cross-referenced this information with a human-readable printout of the strain's genetic program to construct hypothesized mechanism shown in Figure \ref{fig:case_study_1042}.
We hypothesize that cells at the periphery of a registered kin groups send messages backwards over incoming interconnects that induce interconnect-originating cells to send them resource.
Such a mechanism could preferentially increase resource availability at the group periphery, a region where cell-cell conflict is likely elevated.

We performed a series of four-hour competition experiments between wild type and knockout strains to confirm the adaptive nature of each component of this mechanism.
We began by re-routing stimulus 19, which alerts cells to a non-registered-kin neighbor, to activate a known no-op module.
This knockout strain experienced decreased fitness compared to the wild-type strain (16/16 knockout strain extinctions; one-tailed binomial test; $p < 0.0001$; 1996 S.D. 280 cell gens elapsed; \url{https://mmore500.com/hopto/ak}).
Next, we replaced the over-interconnect messaging instruction that triggers over-interconnect resource-sharing with a no-op instruction.
This knockout strain also experienced decreased fitness (16/16 knockout strain extinctions; 1932 S.D. 223 cell gens elapsed; \url{https://mmore500.com/hopto/al}).
We then replaced all eight copies of the over-interconnect resource-sharing instruction triggered by the over-interconnect messaging with no-op instructions, once more yielding a strain with diminished fitness (16/16 knockout strain extinctions; 1860 S.D. 370 cell gens elapsed; \url{https://mmore500.com/hopto/am}).
Finally, we confirmed the soundness of our fitness competition methodology by running control wild-type versus wild-type competitions.
As expected, we observed no effect of strain ID on competition dominance (8/16 knockout strain extinctions; 8/16 wild-type strain extinctions; one-tailed binomial test; $p = 0.60$; 1738 S.D. 217 cell gens elapsed; \url{https://mmore500.com/hopto/aj})

\subsection{Case Study: Interconnect Messaging}

\input{fig/case_study_2032.tex}

This case study was drawn from epoch 18 of batch 32 from the secondary set of evolutionary runs.
We set this strain aside for case study after preliminary screening suggested that over-interconnect messaging played an adaptive role and that the intercellular nature the messaging of was necessary to that adaptation.
The evolutionary history preceeding this case study consumed approximately 72 hours of wall-clock time and 576 compute-core hours.
Approximately 2,197,976 simulation updates and 8,884 cellular generations elapsed.
You can view this case study strain in a live in-browser simulation at \url{https://mmore500.com/hopto/7}.
% 100%|██████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████████| 19/19 [00:00<00:00, 406.39it/s]
% Output saved to title=mastergenerations+_data_hathash_hash=4b3da18cbeb2bc0d+_script_fullcat_hash=5e6b7a64514cbb09+_source_hash=29a51b3-clean+ext=.csv
% level 0
% Generations Elapsed mean 8884.38084673059
% Generations Elapsed std nan
% Generations Elapsed min 8884.38084673059
% Generations Elapsed max 8884.38084673059
% level 1
% Generations Elapsed mean 5674.207208163186
% Generations Elapsed std nan
% Generations Elapsed min 5674.207208163186
% Generations Elapsed max 5674.207208163186
% level 2
% Generations Elapsed mean 1196.6607671890404
% Generations Elapsed std nan
% Generations Elapsed min 1196.6607671890404
% Generations Elapsed max 1196.6607671890404
% EXPIRATION_UPDATE 513408
% EXPIRATION_UPDATE 233216
% EXPIRATION_UPDATE 184096
% EXPIRATION_UPDATE 211488
% EXPIRATION_UPDATE 182368
% EXPIRATION_UPDATE 175840
% EXPIRATION_UPDATE 189952
% EXPIRATION_UPDATE 202496
% EXPIRATION_UPDATE 136992
% EXPIRATION_UPDATE 140576
% EXPIRATION_UPDATE 143712
% EXPIRATION_UPDATE 128512
% EXPIRATION_UPDATE 114048
% EXPIRATION_UPDATE 123136
% EXPIRATION_UPDATE 110176
% EXPIRATION_UPDATE 111296
% EXPIRATION_UPDATE 116832
% EXPIRATION_UPDATE 103264
% EXPIRATION_UPDATE 114208
% EXPIRATION_UPDATE 119136
% EXPIRATION_UPDATE 86560
% EXPIRATION_UPDATE 91104
% EXPIRATION_UPDATE 85984
% EXPIRATION_UPDATE 94688
% EXPIRATION_UPDATE 85856
% EXPIRATION_UPDATE 111200
% EXPIRATION_UPDATE 101536
% EXPIRATION_UPDATE 131456
% EXPIRATION_UPDATE 77440
% EXPIRATION_UPDATE 88352
% EXPIRATION_UPDATE 74944
% EXPIRATION_UPDATE 89280
% EXPIRATION_UPDATE 88512
% EXPIRATION_UPDATE 82240
% EXPIRATION_UPDATE 80576
% EXPIRATION_UPDATE 84192
% EXPIRATION_UPDATE 89344
% EXPIRATION_UPDATE 91104
% EXPIRATION_UPDATE 91456
% EXPIRATION_UPDATE 88384
% EXPIRATION_UPDATE 82656
% EXPIRATION_UPDATE 87296
% EXPIRATION_UPDATE 75200
% EXPIRATION_UPDATE 84992
% EXPIRATION_UPDATE 99328
% EXPIRATION_UPDATE 102176
% EXPIRATION_UPDATE 113568
% EXPIRATION_UPDATE 97760
% EXPIRATION_UPDATE 108864
% EXPIRATION_UPDATE 104736
% EXPIRATION_UPDATE 104416
% EXPIRATION_UPDATE 99872
% EXPIRATION_UPDATE 98880
% EXPIRATION_UPDATE 110176
% EXPIRATION_UPDATE 100224
% EXPIRATION_UPDATE 93440
% EXPIRATION_UPDATE 97888
% EXPIRATION_UPDATE 97952
% EXPIRATION_UPDATE 106880
% EXPIRATION_UPDATE 95776
% EXPIRATION_UPDATE 103424
% EXPIRATION_UPDATE 99584
% EXPIRATION_UPDATE 100576
% EXPIRATION_UPDATE 101216
% EXPIRATION_UPDATE 107904
% EXPIRATION_UPDATE 112608
% EXPIRATION_UPDATE 101408
% EXPIRATION_UPDATE 110976
% EXPIRATION_UPDATE 110560
% EXPIRATION_UPDATE 101472
% EXPIRATION_UPDATE 102336
% EXPIRATION_UPDATE 96864
% EXPIRATION_UPDATE 105472
% EXPIRATION_UPDATE 97664
% EXPIRATION_UPDATE 97280
% EXPIRATION_UPDATE 111520

As before, we began by testing the adaptiveness of the intercellular nature of over-interconnect interaction.
We performed a competition experiment between the wild-type strain and a variant where over-interconnect messages were re-routed back to the sender.\footnote{
This was an independent replication of the initial experiment (performed as part of a wider screen) that singled out the case study strain for further analysis.
}
The wild-type strain was present in greater abundance at the end of all 16 competitions (one-tailed binomial test; $p < 0.0001$; 2/16 variant strain extinctions; 52 S.D. 3 cell gens elapsed).
So the adaptiveness of over-interconnect messaging does depend on the intercellular nature that messaging in this strain.

We proceeded to tease apart the evolved cellular mechanisms this messaging interacts with.
We monitored hardware execution of the wild-type strain in a monoculture population to detect which signals, messages, and fork/call instructions activated each SignalGP module.
Referring to a human-readable printout of the strain's evolved genetic programming, we pieced together the hypothesized mechanism shown in Figure \ref{fig:case_study_2032}.
It appears that neighboring a direct cellular offspring stimulates dispatch of an over-interconnect message that induces the recipient to pause somatic reproduction.

Four-hour competition experiments between wild type and knockout strains allowed us to assess the adaptiveness of each component of this mechanism.
We replaced the instruction responsible for over-interconnect messaging with a no-op instruction and observed a corresponding fitness penalty (16/16 knockout strain extinctions; one-tailed binomial test; $p < 0.0001$; 416 S.D. 58 cell gens elapsed; \url{https://mmore500.com/hopto/aa}).
We also replaced the reproduction-pausing instruction executed in response to over-interconnect messagingg with a no-op.
This caused a similar fitness penalty (16/16 knockout strain extinctions; 401 S.D. 29 cell gens elapsed).

To double-check whether the messaging specifically over interconnects was key to adaptivity we also competed the wild-type strain against variants with the focal over-interconnect messaging instruction substituted for all other possible module-activating instructions:
\begin{itemize}
  \item call (378 S.D. 42 cell gens elapsed; \url{https://mmore500.com/hopto/ac})
  \item fork (377 S.D. 37 cell gens elapsed; \url{https://mmore500.com/hopto/ad})
  \item internal message send (406 S.D 39 cell gens elapsed; \url{https://mmore500.com/hopto/ae})
  \item internal message send-to-all (422 S.D 30 cell gens elapsed; \url{https://mmore500.com/hopto/af})
  \item external message send (377 S.D. 37 cell gens elapsed; \url{https://mmore500.com/hopto/ag})
  \item external message send-to-all (440 S.D. 32 cell gens elapsed; \url{https://mmore500.com/hopto/ah})
\end{itemize}
In each case the substitution variant strain was driven to extinction across all 16 replicate experiments (one-tailed binomial test; $p < 0.0001$).

The directionality of messaging over the interconnect, however, does not appear to affect fitness.
We tried substituting the wild-type instruction, which dispatches a message from the terminus of an interconnect to its origin, with an instruction that instead dispatches a message from the origin of an interconnect to its terminus.
In competition against wild-type, the wild type strain was more abundant in only ten of 16 replicate competitions (one-tailed binomial test; $p = 0.2272$; 14/16 coalesced to a single strain; 410 S.D. 50 cell gen; \url{https://mmore500.com/hopto/ai}).

Satisfied that, indeed, interconnect-transmission of messages contributes to fitness, we set out to characterize the relationship between fitness the distribution of message dispatch and delivery over space and time.
We began by investigating whether non-uniform distribution of message delivery --- and the corresponding reproductive pause activity --- contributed to fitness.
We created a variant strain where instead of activating module 5 and dispatching an over-interconnect message, stimulus 5 directly activates the reproductive pause instructions in module 14.
In this variant strain, we alter stimulus 5 to no longer be triggered by a cell-offspring neighbor.
Instead, we dispatch stimulus 5 with uniform probability based on the empirical module 14 activation rate observed in a monoculture population of the wild-type strain.
% between updates 1000 (238940 hits) and 1100 (267392 hits) updates -> 28452 / (12*45*45*4) = 0.292716
In effect, this manipulation decouples reproductive pause from the distribution of over-interconnect message delivery and instead couples it to a comparable uniform distribution.
Indeed, in competition experiments against the wild-type strain this variant fares poorly (16/16 variant strain extinctions; one-tailed binomial test; $p < 0.0001$; 58 S.D. 2 cell gen).
This suggests that, together, the conjunction of stimulus 5 on a cell and the presence (via a developmental process) of an interconnect to that cell induces a meaningfully (with respect to fitness) non-uniform distribution of message delivery and, subsequently, module 14 activation.

Does the specific distribution of stimulus 5 on message-generating cells contribute to fitness?
As a rough initial exploration of this question, we performed competition experiments with variants where stimulus 5 was always activated, was never activated, and had inverted activation.
In each case, the variant strain was driven to extinction in all 16 replicate runs (one-tailed binomial test; $p < 0.0001$; 48 S.D. 3 cell gens elapsed; 52 S.D. 3 cell gens elapsed; 48 S.D. 3 cell gens elapsed).
Next, we prepared a variant to test if a uniform stimulus distribution with comparable frequency to wild type would exhibit full fitness.
We empirically measured the stimulus 5 activation rate in a monoculture population of the wild-type strain.
%(observe trigger frequency of 22729/(22729+82571) between 1000 and 1100 updates in a monoculture population)
This variant strain also underperformed wild type, being driven to extinction in all 16 replicate competition runs (62 S.D. 3 cell gens elapsed).

It is also possible that the developmental and  process contributes other information to the mix
To randomly select (or select from a distribution).

%TODO Figure \ref{fig:case_study_1042}

