\subsection{Standard SignalGP Instruction Library}

To counteract crowding of the mutational landscape by the volume of custom instructions provided, a second identical copy of each standard SignalGP instruction was included in the library.
These instructions were included in both the neighbor-manager hardware and interconnect-manager hardware instruction libraries.

\begin{itemize}
\item \textbf{Increment}
Increment value in a designated register.
\item \textbf{Decrement}
Decrement value in a designated register.
\item \textbf{Not}
Logically toggle value in a designated register.
\item \textbf{Add}
Add values from two designated registers into a third designated register.
\item \textbf{Subtract}
Subtract values from two designated registers into a third designated register.
\item \textbf{Subtract}
Subtract values from two designated registers into a third designated register.
\item \textbf{Multiply}
Multiply values from two designated registers into a third designated register.
\item \textbf{Divide}
Divide values from two designated registers into a third designated register.
\item \textbf{Modulus}
Calculate the modulus from two designated registers and place result into a third designated register.
\item \textbf{Test Equal}
Compare values in two designated registers and place equality result into a third designated register.
\item \textbf{Test Non-equality}
Compare values in two designated registers and place opposite equality result into a third designated register.
\item \textbf{Test Less}
Compare values in two designated registers and place less-than result into a third designated register.
\item \textbf{If}
If a designated register is non-zero, proceed.
Otherwise, skip block.
\item \textbf{While}
While a designated register is non-zero, loop over a program block.
Otherwise, skip block.
\item \textbf{Countdown}
While a designated register is non-zero, loop over a program block and decrement the value in the designated register.
Otherwise, skip block.
\item \textbf{Close}
If a preceding program block is, close it.
\item \textbf{Break}
Break to the end of the current program block.
\item \textbf{Call}
Call the SignalGP program module that best matches instruction's affinity.
\item \textbf{Return}
If possible, return from the current function.
\item \textbf{Set Memory}
% also covers terminal
Set a designated register's value to genetically-determined value.
\item \textbf{Copy Memory}
Copy the value of a designated register to a second designated register.
\item \textbf{Swap Memory}
Swap the values of two designated registers.
\item \textbf{Input}
Copy a designated element of input memory into a designated register.
\item \textbf{Output}
Copy to a designated element of output memory from a designated register.
\item \textbf{Commit}
Copy a designated register into a designated element of global memory.
\item \textbf{Pull}
Copy a designated element of global memory into a designated register.
\item \textbf{Fork}
Fork a new thread with the SignalGP program module that best matches the instruction's affinity.
\item \textbf{Terminate}
Terminate the current thread.
\item \textbf{Nop}
No operation.
\item \textbf{RNG Draw}
Draw a random value between 0.0 and 1.0 from random number generator and store result in a register.
\item \textbf{Set Regulator}
Set the program module regulator that best matches (without regulation) the instruction's affinity to the value of a designated register.
\item \textbf{Set Own Regulator}
Set the program module regulator of the currently-executing program module to value of a designated register.
\item \textbf{Adjust Regulator}
Adjust the program module regulator of the program module that best matches (without regulation) the instruction's affinity a designated fraction toward a designated register's value.
\item \textbf{Adjust Own Regulator}
Adjust the program module regulator of the currently-executing program module a designated fraction toward a designated register's value.
\item \textbf{Sense Regulator}
Copy the program module regulator value of the program module that best matches (without regulation) the instruction's affinity into a designated register.
\end{itemize}


